---
title: "Clean Analysis"
output: html_document
date: "2023-01-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readxl)
library(ggplot2)
library(PairedData)
library(gridExtra)
library(grid)
library(cowplot)

# SlimStampen
#install.packages("devtools") # Install SlimStampen packages. Instructions on https://github.com/VanRijnLab/SlimStampeRData
library(devtools)
# devtools::install_github("VanRijnLab/SlimStampeRData",
                       # build_vignettes = TRUE,
                       # dependencies = TRUE,
                       # force = TRUE) 
# The console will show a prompt asking whether all packages should be updated. Select option 1 (by typing 1 and then pressing enter), this will update all packages.
vignette("SlimStampeRVignette")
library(SlimStampeRData)
```

Load ADRC data:
```{r}
load("cleandata.Rdata")
lessons <- unique(cleandata$lesson_order)
print(lessons)
textLessons <- cleandata %>% 
  filter(stimcat == "Text") %>% 
  pull(lesson_order) %>% 
  unique()

print(textLessons)
```

Load neurostim data:
```{r, echo=FALSE, results= 'hide'}

# Load data
load("SlimStampen_2022-08-14copy.RData")

# Filter for just neurostim 
groups <- read_csv("groups.csv", show_col_types = FALSE)
data %>% 
  filter(userId %in% groups$userId) -> data

# Calculate fact rep, activation, and alpha 
MAX_ALPHA=0.5
slimdata <- data %>%
  calculate_repetition() %>%
  calculate_alpha_and_activation(maxAlpha=MAX_ALPHA)

```

Clean both data sets to select lessons and variables and order by presentation (repetition)
```{r}
ADRCdata <- cleandata %>% 
  filter(lesson_order == "Swahili 1" |
         lesson_order == "Swahili2" |
         lesson_order == "US Towns 1")  %>% 
  filter(clinicalStatus == "HC") %>% 
  select(userId, factId, repetition, correct) %>% 
  mutate(itemId = paste(userId, factId, sep="-")) %>% 
  group_by(itemId) %>% 
  arrange(repetition, .by_group = TRUE) %>% 
  mutate(type = "ADRC")

NSdata <- slimdata %>% 
  # filter(lessonTitle == "Swahili 1" |
  #       lessonTitle == "European Capitals 1") %>% 
  mutate(userId = as.character(userId)) %>% 
  select(userId, factId, repetition, correct) %>% 
  mutate(itemId = paste(userId, factId, sep="-")) %>% 
  group_by(itemId) %>% 
  arrange(repetition, .by_group = TRUE) %>% 
  mutate(type = "NS")

dataClean <- rbind(ADRCdata, NSdata)

```

Filter the data to select only items where at least 1 error has been committed:
```{r}
errorData <- dataClean %>% 
  mutate(error = case_when(correct == FALSE ~ 1,
                           correct == TRUE ~ 0)) %>% 
  group_by(itemId) %>% 
  summarize(
    totalError = sum(error)
  ) %>% 
  filter(totalError >= 1) %>% 
  left_join(dataClean, by = "itemId")
```

Add in a condition: whether or not an error was made on the 2nd presentation (1 rep) of the item
This is the presentation after the study and immediate test portion so reflects the best opportunity for an early error
```{r}
conditions <- errorData %>% 
  filter(repetition == 1) %>% 
  mutate(condition = case_when(correct == TRUE ~ "No error",
                               correct == FALSE ~ "Error")) %>% 
  select(itemId, condition)
errorCond <- left_join(conditions, errorData, by = "itemId")

```

Filter out people who don't have both conditions:
```{r}
both <- errorCond %>% 
  group_by(userId) %>% 
  summarize(count = n_distinct(condition)) %>% 
  filter(count == 2) %>% 
  pull(userId)

print(both)
```

Make a plot
```{r message=FALSE}
plotData <- errorCond %>% 
  filter(userId %in% both) %>% 
  mutate(correct = case_when(correct == FALSE ~ 0,
                           correct == TRUE ~ 1)) %>% 
  select(userId, condition, repetition, correct) %>% 
  group_by(userId, condition, repetition) %>%
  summarize(
    performance = sum(correct) / n()
  ) %>% 
  mutate("userCond" = paste(userId, condition, sep="-"))

ggplot(plotData) +
  geom_point(mapping = aes(x = repetition, y = performance, color = condition), position = "jitter", alpha = 0.8) + 
  geom_smooth(method= "lm", mapping = aes(x = repetition, y = performance, color = condition, group = userCond), se=F) +
  scale_color_manual(values=c("tomato4","darkseagreen4")) +
  ylab("Accuracy") +
  ggtitle("Overall Performance by Participant") +
  theme(plot.title = element_text(hjust = 0.5, face= "bold"))
```
Break into three portions and test signficance:
```{r}
compareData <- plotData %>% 
  mutate(condition = case_when(condition == "No error" ~ "noError",
                               condition == "Error" ~ "error")) %>% 
  pivot_wider(
    id_cols = c(userId, repetition),
    names_from = condition,
    values_from = performance
  ) %>% 
  filter(repetition != 0 &
         repetition != 1) %>% 
  mutate(difference = error - noError) %>% 
  na.omit()

quality <- compareData %>% 
  group_by(userId) %>% 
  summarize(
    avg = (mean(error) + mean(noError)) / 2
  )

# Because 69425 avg accuracy <50%, we remove them from the data set
compareData <- compareData %>% 
  filter(userId != 69425)

# calculate diff in performance immediately after error
rep2_3 <- compareData %>% 
  filter(repetition %in% c(2,3)) 

test23 <- rep2_3 %>% 
  group_by(userId) %>% 
  summarize(
    error = mean(error),
    noError = mean(noError)
  )

plot23 <- rep2_3 %>% 
  pivot_longer(
    cols = c(error, noError),
    names_to = "condition"
  ) %>% 
  mutate(condition = case_when(condition == "error" ~ "Error",
                               condition == "noError" ~ "No Error")) %>% 
  mutate(repetition = as.factor(repetition)) %>% 
  mutate(userCond = paste(userId, condition, sep="-"))

rep4_6 <- compareData %>% 
  filter(repetition %in% c(4,5, 6)) 

test46 <- rep4_6 %>% 
  group_by(userId) %>% 
  summarize(
    error = mean(error),
    noError = mean(noError)
  )

plot46 <- rep4_6 %>% 
  pivot_longer(
    cols = c(error, noError),
    names_to = "condition"
  ) %>% 
  mutate(condition = case_when(condition == "error" ~ "Error",
                               condition == "noError" ~ "No Error")) %>% 
  mutate(repetition = as.factor(repetition)) %>% 
  mutate(userCond = paste(userId, condition, sep="-"))

rep7_max <- compareData %>% 
  filter(repetition > 6) 

test7max <- rep7_max %>% 
  group_by(userId) %>% 
  summarize(
    error = mean(error),
    noError = mean(noError)
  )

plot7max <- rep7_max %>% 
  pivot_longer(
    cols = c(error, noError),
    names_to = "condition"
  ) %>% 
  mutate(condition = case_when(condition == "error" ~ "Error",
                               condition == "noError" ~ "No Error")) %>% 
  mutate(repetition = as.factor(repetition)) %>% 
  mutate(userCond = paste(userId, condition, sep="-"))

```

```{r message=FALSE}
plot2_3A <- ggplot(plot23, alpha = 0.8) +
  geom_point(mapping = aes(x = repetition, y = value, color = condition, group = userId)) +
  geom_smooth(method= "lm", mapping = aes(x = repetition, y = value, color = condition, group = userCond), se=F) +
  labs(y = "Accuracy", x = "Fact Presentation") +
  scale_y_continuous(limits=c(0, 1.1), breaks=c(0, 0.25, 0.5, 0.75, 1)) +
  scale_color_manual(values=c("tomato4","darkseagreen4"))

plot4_6A <- ggplot(plot46, alpha = 0.8) +
  geom_point(mapping = aes(x = repetition, y = value, color = condition, group = userId)) +
  geom_smooth(method= "lm", mapping = aes(x = repetition, y = value, color = condition, group = userCond), se=F) +
  labs(y = "Accuracy", x = "Fact Presentation") +
  scale_y_continuous(limits=c(0, 1.1), breaks=c(0, 0.25, 0.5, 0.75, 1)) +
  theme(legend.position="none") +
  scale_color_manual(values=c("tomato4","darkseagreen4"))

plot7_maxA <- ggplot(plot7max, alpha = 0.8) +
  geom_point(mapping = aes(x = repetition, y = value, color = condition, group = userId)) +
  geom_smooth(method= "lm", mapping = aes(x = repetition, y = value, color = condition, group = userCond), se=F) +
  labs(y = "Accuracy", x = "Fact Presentation") +
  scale_y_continuous(limits=c(0, 1.1), breaks=c(0, 0.25, 0.5, 0.75, 1)) +
  theme(legend.position="none") +
  scale_color_manual(values=c("tomato4","darkseagreen4"))

# save legend
get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

legend <- get_legend(plot2_3A)
plot2_3A <- plot2_3A + theme(legend.position="none")

grid.arrange(plot2_3A, plot4_6A, plot7_maxA, legend, ncol=2, top=textGrob("Performance Across Repetitions"))

```

Stats test for Reps 2-3
```{r}
# Rep 2-3
error23 <- test23$error
noError23 <- test23$noError
pd23 <- paired(error23, noError23)
p23 <- plot(pd23, main = "Repetitions 2-3", type = "profile") + theme_bw()
p23


box23 <- test23 %>% 
  pivot_longer(
    cols = c(error, noError),
    names_to = "condition"
  )

meanStd23 <- box23 %>% 
  group_by(condition) %>% 
  summarize(
    average = mean(value),
    std = sd(value)
  )

p23 <- ggplot(box23, aes(x = condition, y = value, color = condition)) +
  geom_boxplot() +
  ggtitle("(A) Repetitions 2-3") +
  ylab("") + 
  scale_color_manual(values=c("tomato4","darkseagreen4")) +
  theme(plot.title = element_text(hjust = 0, size=10))
p23

legend2 = get_legend(p23)
p23 <- p23 + theme(legend.position = "none")

result23 <- t.test(error23, noError23, paired =TRUE)
result23
```

Stats test for Reps 4-6
```{r}
?plot

error46 <- test46$error
noError46 <- test46$noError
pd46 <- paired(error46, noError46)
p46 <- plot(pd46, type = "profile") + theme_bw()
p46

box46 <- test46 %>% 
  pivot_longer(
    cols = c(error, noError),
    names_to = "condition"
  )

meanStd46 <- box46 %>% 
  group_by(condition) %>% 
  summarize(
    average = mean(value),
    std = sd(value)
  )

p46 <- ggplot(box46, aes(x = condition, y = value, color = condition)) +
  geom_boxplot() +
  ggtitle("(B) Repetitions 4-6") +
  ylab("") + 
  scale_color_manual(values=c("tomato4","darkseagreen4")) +
  theme(plot.title = element_text(hjust = 0, size=10), legend.position = "none")
p46

result46 <- t.test(error46, noError46, paired =TRUE)
result46
```

Stats test for Reps 7max
```{r}
error7max <- test7max$error
noError7max <- test7max$noError
pd7max <- paired(error7max, noError7max)
p7max <- plot(pd7max, type = "profile") + theme_bw()
p7max

box7max <- test7max %>% 
  pivot_longer(
    cols = c(error, noError),
    names_to = "condition"
  )

meanStd7max <- box7max %>% 
  group_by(condition) %>% 
  summarize(
    average = mean(value),
    std = sd(value)
  )

p7max <- ggplot(box7max, aes(x = condition, y = value, color = condition)) +
  geom_boxplot() +
  ggtitle("(C) Repetitions 7+") +
  ylab("") + 
  scale_color_manual(values=c("tomato4","darkseagreen4")) +
  theme(plot.title = element_text(hjust = 0, size=10), legend.position = "none")
p7max

result7max <- t.test(error7max, noError7max, paired =TRUE)
result7max
```


Group paired plots:
```{r}
grid.arrange(p23, p46, p7max, legend2, nrow = 2,
                         top = textGrob("Paired Comparisons of Conditions Across Testing Phases", vjust = 1, gp = gpar(fontface = "bold", cex = 1)),
                         left = textGrob("Accuracy", rot = 90, vjust = 1))

```
